Stability of variational quantum dynamics with Gaussian approximants

### 2013-07-17 paper.tex
writing

@@

The invention of the laser called for a quantum theory of intense radiation.  Laser beams are difficult because their quantum descriptions are so large: they involve up to $10^{18}$ photons, distributed across thousands of modes.  The theory was made managable by Glauber and Sudarshan, who showed that any quantum state can be expanded as a superposition of coherent states.  Laser light is very coherent, so these expansions often represent it efficiently.  Moyal(?) developed an efficient way to calculate dynamics, leading to a family of phase space Monte-Carlo methods.  When atomic Bose-Einstein condensates were made, and very coherent matter became easy to access in the lab, phase space methods were sucessfully applied to calculate its dynamics, and have since been applied to other forms of condensed matter.  They have solved many problems, but are limited by intractable convergence issues.

@@

When used on computers, the variational methods expand quantum states over finitely many coherent states.  This is a discretisation of the expansion introduced by Glauber, a mapping $G$ from square-integrable functions to Fock space, where
$$Gf={1\over π}\int g(α*)|α〉\,d²α.\eqno{(1)}$$
This is different from Glauber's use: he uses $f=ge^{½|α|²}$.  Is it easier to work Glauber's way, with the $e^{-½|α|²}$ in the mapping, so that the singular vectors are entire functions?  You'd have to use a Bargmann inner product to normalise them; the theory of that is well developed.

@@

>>>>

\def\G{|Γ〉}
\def\cite#1{$[\hbox{\tt #1}]$}

@@

The invention of the laser called for a quantum theory of intense radiation.  It is difficult to analyse laser beams because their quantum descriptions are so large, comprising up to $10^{18}$ photons distributed across thousands of modes.  The theory was made managable by Glauber and Sudarshan, who showed that any quantum state can be expanded as a superposition of coherent states.  Laser light is very coherent, so these expansions often represent it efficiently.  Moyal(?) developed an efficient way to calculate dynamics, leading to a family of phase space Monte-Carlo methods.  When atomic Bose-Einstein condensates were made, and very coherent matter became easy to access in the lab, phase space methods were sucessfully applied to calculate its dynamics, and have since been applied to other forms of condensed matter.  They have solved many problems, but are limited by intractable convergence issues.

@@

@@

The expansion operator introduced by Glauber\cite{prx-131-2766} is
$$\G f={1\over π}\int |α〉e^{-½|α|²}f(α*)\,d²α.$$
The notation $\int f(α)\,d²α$ means $\int_{-∞}^{∞}\int_{-∞}^{∞} f(x+iy)\,dx\,dy$.  The function $f$ is a function of a complex variable, such that $e^{-½|α|²}f(α)$ is square-integrable.  Bargmann\cite{} showed that these functions form a Hilbert space under the inner product $〈g,f〉=\int e^{-|α|²}g*(α)f(α)\,d²α$; we will call it the space of phase space distributions, or just distributions.  In this paper, the application of a linear function is written as a product, and functions whose range is ket space are denoted by kets, usually with a greek letter.  So $\G f$ is the ket that results from applying the linear operator $\G$ to the phase space distribution $f$.

The state $|α〉$ is a coherent state with complex amplitude $α$.  These can be defined in a number of ways, perhaps most simply in second quantisation, as eigenstates of a boson annihilation operator, with $a|α〉=α|α〉$.  The annihilation operator is not hermitian, so the eigenvalues $α$ are complex numbers.  In first quantisation, such an operator is the complex amplitude
$$a=?(λx+ip/λ),\qquad\hbox{Get constants from Cahill}$$
where $x$ and $p$ are the position and momentum operators; the bosons in this case are vibrational quanta in a notional harmonic potential, centred at the origin, with frequency ?.

Coherent states are not orthogonal, but their brackets follow the law
$$〈β|α〉=e^{-½|β|²-½|α|²+β*α},$$
whence
$$|〈β|α〉|²=e^{-|β-α|²}.$$
Their projector follows
$$\int |α〉〈α|\,d²α=π.$$
For an orthonormal basis, the right hand side would be 1 instead of $π$.  It follows that the operator $\G$ is onto: let $|ψ〉$ any ket, and set $f₀(α*)=e^{½|α|²}〈α|ψ〉$.  Then
$$\G f₀={1\over π}\int |α〉〈α|ψ〉\,d²α=|ψ〉.$$
Furthermore, coherent states have the property that $e^{½|α|²}|α〉$ is an entire function of $α$, so $f₀$ is an entire function.  Glauber showed that it is the unique entire solution to $\G f₀=|ψ〉$.  Further properties of the coherent states will be stated as required.  Glauber's summary of them is paraphrased in every quantum optics textbook ever written.

Obviously, $\G$ is not one to one, because the images of the entire functions already fill ket space.
